---
title: "Global Greenhouse Gas (GHG) Emissions Report"
output: pdf_document
header-includes:
  - \usepackage{titling} 
  - \setlength{\droptitle}{-2.5cm} 
  - \posttitle{\end{center}\vspace{-2cm}} 
---

\begin{center}
\normalsize{\textit{Giulia Bertoldo}}
\end{center}

```{r setup, include=FALSE}
# Set up and variables 
knitr::opts_chunk$set(echo = TRUE)
path = 'data/EDGAR_2024_GHG_booklet_2024.xlsx'
# Load packages
library(tidyverse)
library(readxl)
library(purrr)
library(stringr)
```

## Introduction 
The report presents an analysis of greenhouse gas (GHG) emissions data from the Emissions Database for Global Atmospheric Research (EDGAR) [1]. Data cover all countries from 1970 to 2023. 

Data were analyzed and visualized using R, mainly with the tidyverse packages (e.g., dplyr, ggplot2). The report was generated using Rmarkdown, and Git and GitHub were used as version control system (\href{https://github.com/giuliabertoldo/GHG-Emissions}{\textcolor{blue}{link to GitHub repository}}) .

## Evolution of GHG emission growth

*What is the evolution of GHG growth in the Euro area, in the European Union, and worldwide?*

For this analysis, data from the Excel sheet 'GHG_totals_by_country' were used. The pre-existing GLOBAL TOTAL and EU27 rows were used, while new rows for Eurozone were computed aggregating data from the relevant countries. 

* From 1970 to 2023, worldwide GHG emissions increased from approximately 24000 Mt CO2eq/yr in 1970 to about 52000 Mt CO2eq/yr in 2023.

* Throughout this period, the European Union consistently contributed around 5,000 Mt CO2eq/yr until 1990. After that, emissions in the European Union began to decline, with a more significant drop following 2008.

* The Eurozone followed a similar declining trend to the European Union after 2008. 

```{r include=FALSE}
# Read data
df_ghg_totals_by_country <- read_excel(path, sheet = 'GHG_totals_by_country')

# DATA QUALITY -------------------------------------------------------------------------------
#### Data types checks ####
glimpse(df_ghg_totals_by_country)

# `EDGAR Country Code`
df_ghg_totals_by_country$`EDGAR Country Code`
df_ghg_totals_by_country$Country

#### Upper case country name ####
df_ghg_totals_by_country <- df_ghg_totals_by_country %>%
  mutate (Country = str_to_upper(Country))

####  Missing values ####
as.data.frame(map(df_ghg_totals_by_country, ~sum(is.na(.))))

df_na <- df_ghg_totals_by_country %>%
  filter(if_any(everything(), is.na))

# Filter out two rows with missing values
df_ghg_totals_by_country <- df_ghg_totals_by_country %>%
  filter(if_all(everything(), ~ !is.na(.)))

####  Duplicates ####
duplicate_rows <- df_ghg_totals_by_country %>%
  duplicated(.) | duplicated(df_ghg_totals_by_country, fromLast = TRUE)
sum(duplicate_rows)

#### Factors####
# Convert character columns to factor
df_ghg_totals_by_country <- df_ghg_totals_by_country %>%
  mutate(across(where(is.character), factor))

#### EUROZONE FLAG, EU27 FLAG, GLOBAL FLAG ####
# eurozone
eurozone <- c('AUSTRIA', 'BELGIUM', 'CROATIA', 'CYPRUS', 'ESTONIA', 'FINLAND',
              'FRANCE AND MONACO', 'GERMANY', 'GREECE', 'IRELAND',
              'ITALY, SAN MARINO AND THE HOLY SEE', 'LATVIA', 'LITHUANIA',
              'LUXEMBOURG', 'MALTA', 'NETHERLANDS', 'PORTUGAL', 'SLOVAKIA',
              'SLOVENIA', 'SPAIN AND ANDORRA')

# Keep only relevant rows
df_ghg_totals_by_country <- df_ghg_totals_by_country %>%
  filter((Country %in% c('EU27', 'GLOBAL TOTAL')) | (Country %in% eurozone))

df_ghg_totals_by_country <- df_ghg_totals_by_country %>%
  mutate(GEO_TYPE = case_when(
    Country == 'EU27' ~ 'EU27',
    Country == 'GLOBAL TOTAL' ~ 'GLOBAL',
    Country %in% eurozone ~ 'EUROZONE'
  ))

check_eurozone <- df_ghg_totals_by_country %>%
  select(Country, GEO_TYPE)
print(check_eurozone, n = Inf )

#### Pivot longer ####
df_ghg_totals_by_country_long <- df_ghg_totals_by_country %>%
  pivot_longer(cols = `1970`: `2023`, names_to = 'YEAR', values_to = 'GHG_EMM')

glimpse(df_ghg_totals_by_country_long)

#### Select column ####
df_ghg_totals_by_country_long <- df_ghg_totals_by_country_long %>%
  select(GEO_TYPE, YEAR, GHG_EMM)

#### Aggregate ####
df_ghg_totals_by_country_group <- df_ghg_totals_by_country_long %>%
  group_by(GEO_TYPE, YEAR) %>%
  summarize(TOT_GHG_EMM = sum(GHG_EMM), .groups = "drop")
glimpse(df_ghg_totals_by_country_group)

#### Geography Type as factor
df_ghg_totals_by_country_group$GEO_TYPE <- factor(df_ghg_totals_by_country_group$GEO_TYPE,
                                                  levels = c('GLOBAL',
                                                             'EU27',
                                                             'EUROZONE'),
                                                  labels = c("Worldwide",
                                                             "European Union",
                                                             "Eurozone")
)


```

```{r echo = FALSE}
#### CHART 1 --------------------------------------------------
custom_colors <- c("#40004B", "#1B7837", "#A6DBA0")

ggplot(df_ghg_totals_by_country_group, aes(x = YEAR, y = TOT_GHG_EMM, color = factor(GEO_TYPE), fill = factor(GEO_TYPE), group = GEO_TYPE)) +
  geom_line() +  
  geom_ribbon(alpha = 0.4, aes(ymin = 0, ymax = TOT_GHG_EMM))+  
  labs(
    title = 'Evolution of GHG Emissions',
    x = 'Year',
    y = 'GHG (Mt CO2eq/yr)',
    color = 'Geography'
  ) +
  scale_color_manual(values = custom_colors) +  # Use your custom color palette
  scale_fill_manual(values = custom_colors) +   # Ensure fill uses the same colors as the lines
  theme_light() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 10),
    legend.key.size = unit(0.3, "cm"),
    legend.spacing.y = unit(0.2, "cm"),
    legend.position = "bottom"
  ) +
  guides(fill = "none")  +
  scale_x_discrete(breaks = df_ghg_totals_by_country_group$YEAR[seq(1, length(df_ghg_totals_by_country_group$YEAR), by = 2)]) +
  scale_y_continuous(breaks = seq(0, max(df_ghg_totals_by_country_group$TOT_GHG_EMM), by = 5000))

```

# Countries GHG emission comparison

*How do countries (classified as in The World Bank income groups) compare with respect to GHG emissions per capita?*

For this analysis, data from the Excel sheet 'GHG_per_capita_by_country' were used. The country classification by income group was sourced from \href{https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups?}{\textcolor{blue}{The World Bank Data Help Desk}}.

* Between 1970 and 1980, high income countries emitted more than twice the amount of GHG per capita compared to the combined total of upper middle income, lower middle income and low income countries.  

* From 1970 to 2023, high income countries displayed an overall decrease in GHG emissions per capita, while the other countries showed a stable trend. 

```{r include=FALSE}
# Read data
path = 'data/EDGAR_2024_GHG_booklet_2024.xlsx'
df_per_capita_by_country <- read_excel(path, sheet = 'GHG_per_capita_by_country')
df_class <- read_excel('data/CLASS.xlsx')
# DATA QUALITY -------------------------------------------------------------------------------
#### Data types checks ####
glimpse(df_per_capita_by_country)
glimpse(df_class)

#### Upper case country name ####
df_per_capita_by_country <- df_per_capita_by_country %>%
  mutate (Country = str_to_upper(Country))
df_per_capita_by_country$Country

df_class <- df_class %>%
  mutate (Economy = str_to_upper(Economy))
df_class$Economy

####  Missing values ####
# GHG data
as.data.frame(map(df_per_capita_by_country, ~sum(is.na(.))))
df_na <- df_per_capita_by_country %>%
  filter(if_any(everything(), is.na))
# Filter out two rows with missing values
df_per_capita_by_country <- df_per_capita_by_country %>%
  filter(if_all(everything(), ~ !is.na(.)))

# Income data
as.data.frame(map(df_class, ~sum(is.na(.))))
df_na_ingro <- df_class %>%
  filter(is.na(Economy) == TRUE)
df_na_ingro
# Filter out row with missing values
df_class <- df_class %>%
  filter(is.na(Economy) != TRUE)
# Venezuela is not classified into an income group + other regional areas
null_income <- df_class %>%
  filter(is.na(`Income group`) == TRUE)
print(null_income, n = Inf)

####  Duplicates ####
duplicate_rows <- df_per_capita_by_country %>%
  duplicated(.) | duplicated(df_per_capita_by_country, fromLast = TRUE)
sum(duplicate_rows)

#### Factors ####
# Convert character columns to factor
df_per_capita_by_country <- df_per_capita_by_country %>%
  mutate(across(where(is.character), factor))

#### Left Outer join ####
df_per_capita_by_country_with_income <- left_join(df_per_capita_by_country,
                                                  df_class %>%
                                                    select(Code,`Income group`),
                                                  by = c('EDGAR Country Code' = 'Code'))

# Check join results
glimpse(df_per_capita_by_country_with_income)

# Apart from Venezuela, EU27, GLOBAL, other 12 countries have not been assigned an income group
df_per_capita_by_country_with_income %>%
  filter(is.na(`Income group`)) %>%
  select(`EDGAR Country Code`, 'Country', `Income group` )

# Drop rows without income group
df_per_capita_by_country_with_income <- df_per_capita_by_country_with_income %>%
  filter(is.na(`Income group`) == FALSE)

#### Pivot longer ####
df_per_capita_by_income_long <- df_per_capita_by_country_with_income %>%
  pivot_longer(cols = `1970`: `2023`, names_to = 'YEAR', values_to = 'GHG_EMM')

glimpse(df_per_capita_by_income_long)

#### Aggregate ####
df_per_capita_by_income_long_gr <- df_per_capita_by_income_long %>%
  group_by(`Income group`, YEAR) %>%
  summarize(TOT_GHG_EMM = sum(GHG_EMM), .groups = "drop")
df_per_capita_by_income_long_gr

#### Income groups as factos ####
df_per_capita_by_income_long_gr$`Income group` <- factor(df_per_capita_by_income_long_gr$`Income group`,
                                                         levels = c("High income",
                                                                    "Upper middle income",
                                                                    "Lower middle income",
                                                                    "Low income")
)

glimpse(df_per_capita_by_income_long_gr)
```


```{r echo = FALSE}
#### CHART 2 --------------------------------------------------
## Choose colors
custom_color <- c("#FEE0D2", "#FCBBA1", "#FB6A4A", "#BD0026")
## Stacked bar chart
ggplot(df_per_capita_by_income_long_gr, aes(x = YEAR,
                                            y = TOT_GHG_EMM,
                                            fill = factor(`Income group`),
                                            group = `Income group`)) +
  geom_bar(stat = 'identity', position = 'stack',color = "white", linewidth = 0.5) +
  scale_fill_manual(values = c("High income" =  "#BD0026",
                               "Upper middle income" = "#FB6A4A",
                               "Lower middle income" = "#FCBBA1",
                               "Low income" ="#FEE0D2")) +
  labs(
    title = 'Country (Income Group) Contribution to GHG Emissions',
    x = 'Year',
    y = 'GHG (t CO2eq/cap/yr)',
    fill = 'Income Group'
  ) +
  theme_light() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 10),
    legend.key.size = unit(0.3, "cm"),
    legend.spacing.y = unit(0.2, "cm"),
    legend.position = "bottom"
  ) +
  scale_x_discrete(breaks = df_per_capita_by_income_long_gr$YEAR[seq(1, length(df_per_capita_by_income_long_gr$YEAR), by = 2)]) +
  scale_y_continuous(breaks = seq(0, 2000, by = 250))

```

\newpage

# Contribution of individual countries and continents to worlwide GHG emissions

*What is the contribution of individual countries and continents to the worldwide GHG emissions?*

For this analysis, data from the Excel sheet 'GHG_totals_by_country' were used. Emissions data were summed across years, and the percentage of total GHG emissions was calculated for each country. Countries were classified into continents using data from the *countrycode* R package. 

* China (19%) and United States (17.5%) are the leading contributors to global GHG emissions, followed by Russia (6.45%) and India (5.38%).

* Asia is the continent with the largest share of global GHG emissions, accounting for nearly 45%, while Oceania has the smallest contribution.

* The Americas rank second, contributing just over 25% of global emissions.

```{r include= FALSE}
# Load library 
library(countrycode)
library(maps)
library(patchwork)
# Read data
df_ghg_totals_by_country <- read_excel(path, sheet = 'GHG_totals_by_country')


# DATA QUALITY  (common with task 1 )-------------------------------------------------------------------------------
#### Data types checks ####
glimpse(df_ghg_totals_by_country)

# `EDGAR Country Code`
df_ghg_totals_by_country$`EDGAR Country Code`
df_ghg_totals_by_country$Country

#### Upper case country name ####
df_ghg_totals_by_country <- df_ghg_totals_by_country %>%
  mutate (Country = str_to_upper(Country))

####  Missing values ####
df_na <- df_ghg_totals_by_country %>%
  filter(if_any(everything(), is.na))
df_na

# Filter out rows with missing values
df_ghg_totals_by_country <- df_ghg_totals_by_country %>%
  filter(if_all(everything(), ~ !is.na(.)))

####  Duplicates ####
duplicate_rows <- df_ghg_totals_by_country %>%
  duplicated(.) | duplicated(df_ghg_totals_by_country, fromLast = TRUE)
sum(duplicate_rows)

#### Exclude EU27, GLOBAL TOTAL, INTERNATIONAL AVIATION, INTERNATIONAL SHIPPING ####
df_ghg_totals_by_country <- df_ghg_totals_by_country %>%
  filter(!(Country %in% c('EU27', 'GLOBAL TOTAL', 'INTERNATIONAL AVIATION', 'INTERNATIONAL SHIPPING' )))

#### Pivot longer ####
df_ghg_totals_by_country_long <- df_ghg_totals_by_country %>%
  pivot_longer(cols = `1970`: `2023`, names_to = 'YEAR', values_to = 'GHG_EMM')
glimpse(df_ghg_totals_by_country_long)

#### Group by Country and sum across years ####
df_ghg_totals_by_country_long_gr <- df_ghg_totals_by_country_long %>%
  group_by(Country, `EDGAR Country Code`) %>%
  summarize(TOT_GHG_EMM = sum(GHG_EMM), .groups = "drop")
glimpse(df_ghg_totals_by_country_long_gr)

#### Add percentage total ####
grand_tot_ghg <- sum(df_ghg_totals_by_country_long_gr$TOT_GHG_EMM)

df_ghg_perc_by_country <- df_ghg_totals_by_country_long_gr %>%
  mutate(PER_GHG_EMM = round((TOT_GHG_EMM/grand_tot_ghg)*100, 4))
glimpse(df_ghg_perc_by_country)

#### Categorize percentage ####
min(df_ghg_perc_by_country$PER_GHG_EMM)
max(df_ghg_perc_by_country$PER_GHG_EMM)

df_ghg_perc_by_country$CAT_PER_GHG_EMM <- cut(
  df_ghg_perc_by_country$PER_GHG_EMM,
  breaks = seq(0, 20, by = 2.5),       
  include.lowest = TRUE,               
  right = FALSE,                       
  labels = paste0(seq(0, 17.5, by = 2.5), "% - ", seq(2.5, 20, by = 2.5), "%")
)
glimpse(df_ghg_perc_by_country)

#### Add continent ####
codes <- countrycode::codelist
country_continet <- codes %>%
  select(iso3c, continent)

# Missing values
df_na <- country_continet %>%
  filter(if_any(everything(), is.na))
df_na

# Filter rows with missing values
country_continent <- country_continet %>%
  filter(if_all(everything(), ~ !is.na(.)))

glimpse(country_continent)

# Left join
df_ghg_perc_by_country_cont <- left_join(df_ghg_perc_by_country,
                                         country_continent,
                                         by = c('EDGAR Country Code' = 'iso3c' ))
glimpse(df_ghg_perc_by_country_cont)

# Missing values
df_ghg_perc_by_country_cont %>%
  filter(is.na(continent) == TRUE)

df_ghg_perc_by_country_cont$continent[df_ghg_perc_by_country_cont$`EDGAR Country Code` == 'ANT'] <- "Americas"
df_ghg_perc_by_country_cont$continent[df_ghg_perc_by_country_cont$`EDGAR Country Code` == 'SCG'] <- "Europe"

glimpse(df_ghg_perc_by_country_cont)

# Table with continent data
continent <- df_ghg_perc_by_country_cont %>%
  select(continent, PER_GHG_EMM) %>%
  group_by(continent)%>%
  summarize(CONT_PER_GHG_EMM = sum(PER_GHG_EMM)) %>%
  arrange(desc(CONT_PER_GHG_EMM))

#### Worlds map data Prep ####
world <- map_data("world")

# Upper case country name
world <- world %>%
  mutate(region = str_to_upper(region))

# Align dataset name with maps data requirements (no corrispondence found for GIBRALTAR, HONG KONG, MACAO, CONGO)
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'ANTIGUA AND BARBUDA'] <- 'ANTIGUA'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'BRITISH VIRGIN ISLANDS'] <- 'VIRGIN ISLANDS'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'CABO VERDE'] <- 'CAPE VERDE'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'CÔTE D’IVOIRE'] <- 'IVORY COAST'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'CURAÇAO'] <- 'CURACAO'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'CZECHIA'] <- 'CZECH REPUBLIC'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'ESWATINI'] <- 'SWAZILAND'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'FAROES'] <- 'FAROE ISLANDS'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'FRANCE AND MONACO'] <- 'FRANCE'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'FRANCE AND MONACO'] <- 'MONACO'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'ISRAEL AND PALESTINE, STATE OF'] <- 'ISRAEL'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'ITALY, SAN MARINO AND THE HOLY SEE'] <- 'ITALY'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'MYANMAR/BURMA'] <- 'MYANMAR'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'RÉUNION'] <- 'REUNION'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'SAINT KITTS AND NEVIS'] <- 'SAINT KITTS'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'SAINT VINCENT AND THE GRENADINES'] <- 'SAINT VINCENT'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'SÃO TOMÉ AND PRÍNCIPE'] <- 'SAO TOME AND PRINCIPE'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'SERBIA AND MONTENEGRO'] <- 'SERBIA'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'SERBIA AND MONTENEGRO'] <- 'MONTENEGRO'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'SPAIN AND ANDORRA'] <- 'SPAIN'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'SPAIN AND ANDORRA'] <- 'ANDORRA'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'SUDAN AND SOUTH SUDAN'] <- 'SUDAN'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'SWITZERLAND AND LIECHTENSTEIN'] <- 'SWITZERLAND'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'SWITZERLAND AND LIECHTENSTEIN'] <- 'LIECHTENSTEIN'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'THE GAMBIA'] <- 'GAMBIA'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'TRINIDAD AND TOBAGO'] <- 'TRINIDAD'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'TRINIDAD AND TOBAGO'] <- 'TOBAGO'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'TÜRKIYE'] <- 'TURKEY'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'UNITED STATES'] <- 'USA'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'VIET NAM'] <- 'VIETNAM'
df_ghg_perc_by_country_cont$Country[df_ghg_perc_by_country_cont$Country == 'UNITED KINGDOM'] <- 'UK'

# Left join
df_ghg_perc_by_country_world <- left_join(world,
                                          df_ghg_perc_by_country_cont,
                                          by = c('region' = 'Country' ))
glimpse(df_ghg_perc_by_country_world)

#Find unmapped countries
codes_in <- df_ghg_perc_by_country_world %>%
  select (`EDGAR Country Code`) %>%
  distinct()

codes_out <- df_ghg_perc_by_country_cont %>%
  select(`EDGAR Country Code`, 'Country') %>%
  distinct() %>%
  filter(!(`EDGAR Country Code` %in% codes_in$`EDGAR Country Code`))
print(codes_out, n = Inf)

# Complete list of regions
regions <- world %>%
  select (region) %>%
  distinct() %>%
  arrange(region)
```




```{r echo=FALSE, warning=FALSE}
# VISUALIZATION WORLD MAP --------------------------------------
map_world <- ggplot() +
  geom_map(data=df_ghg_perc_by_country_world,
           map=world,
           aes(map_id=region, x=long, y=lat, fill=CAT_PER_GHG_EMM), color = "black", linewidth = 0.1) +
  scale_fill_brewer(palette = "Reds",
                    name = "GHG Emissions Contribution")+
  coord_equal() +
  theme_light() +
  theme(
    axis.title = element_blank(),  
    axis.text = element_blank(),   
    axis.ticks = element_blank(),  
    legend.position = "bottom"     
  )

# CONTINENTS BARCHARTS -----------------------------------------------------------------------------
continent_bar <- ggplot(data = continent, aes(x = reorder(continent, CONT_PER_GHG_EMM), y = CONT_PER_GHG_EMM)) +
  geom_bar(stat = "identity", width = 0.5, position = position_dodge(width = 0.5),  fill = "lightgrey", color = "darkgrey") +
  labs(
    x = 'Continent',
    y = 'Percentage Contribution to Worldwide GHG'
  ) +
  theme_light() +
  theme(
    legend.position = "none",
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 12, hjust = 1)
  ) +
  coord_flip() + 
  scale_y_continuous(breaks = seq(0, 50, by = 5))

# COMBINE GRAPHS -----------------------------------------------------------------------------
# Combine the map and bar chart using patchwork
combined_plot <- map_world / continent_bar + plot_layout(heights = c(2, 1))

# Display the combined plot
combined_plot
```


# References  

[1] EDGAR (Emissions Database for Global Atmospheric Research) Community GHG Database (a collaboration between the European Commission, Joint Research Centre (JRC), the International Energy Agency (IEA), and comprising IEA-EDGAR CO2, EDGAR CH4, EDGAR N2O, EDGAR F-GASES version EDGAR_2024_GHG (2024) European Commission
